# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
So this assumes that the file activity.csv is in the main folder.
Then the file can be loaded by using:
```{r}
data <- read.csv("activity.csv")
```
Quick inspection of the data shows that date is a Factor. A conversion to date
objects follows.
```{r}
str(data)
data[,2] <- as.Date(data[,2])
```

## What is mean total number of steps taken per day?
To calculate this I will use the aggregate function and then make a histogram.
Both mean and median are calculated and all NA are removed first.
```{r}
data1 <- na.omit(data)
calculated_mean <- aggregate(steps~date, data=data1,FUN=mean)
calculated_median <- aggregate(steps~date, data=data1,FUN=median)
hist(calculated_mean$steps)
calculated_mean
calculated_median
```
So apparently the median total number of steps is 0 for all days due to there
being more zeroes than anything else in this dataset.

## What is the average daily activity pattern?
To find the average daily pattern the aggregate is taken for the steps and
interval instead of the date. This data is plotted.
```{r}
averaged_mean <- aggregate(steps~interval, data=data1,FUN=mean)
plot(averaged_mean, type="l")
```
The maximum interval can be found as:
```{r}
averaged_mean[which(averaged_mean$interval==max(averaged_mean$interval)),1]
```


## Imputing missing values
So since I removed the na values earlier the number of nas can be found as
taking the diffence in rows from the untouched data.
```{r}
nrow(data)-nrow(data1)
```

The missing values will be filled in the with averaged rounded-off step count
for the interval.
```{r}
data2 <- data
Sys.setlocale("LC_TIME","English")
for(i in 1:nrow(data2))
    {
    if(is.na(data2[i,1]))
        {
        data2[i,1] <- round(averaged_mean[which(averaged_mean[,1]==data2[i,3]),2],digits=0)
        }
    }
```
So to compare to the dataset with the removed values.
```{r}
calculated_mean2 <- aggregate(steps~date, data=data2,FUN=mean)
calculated_median2 <- aggregate(steps~date, data=data2,FUN=median)
hist(calculated_mean2$steps)
calculated_mean
calculated_mean2
calculated_median
calculated_median2
```
So there are some differences mostly with regards to median where some days
have medians that are non-zero compared to the others with medians with zeroes.
The mean has also changed depending on those days that have the most nas.

## Are there differences in activity patterns between weekdays and weekends?
First I need to add a new column with data on weekdays or weekends.
NOTE: this is done on a Danish computer. If this needs to run on other machines
then change the statements to read to your language. 
(Extra note: I cannot change locale for some reason)
```{r}
data2$weekday <- weekdays(data2[,2])
for(i in 1:nrow(data2))
    {
    if(data2[i,4] %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
        {
        data2[i,4] <- "weekday"
        }
    else
        {
        data2[i,4] <- "weekend"
        }
    }
```

Now for generating the two plots. First the dataset is divided into weekday and
weekend data sets.
```{r}
weekday <- data2[which(data2[,4]=="weekday"),]
weekend <- data2[which(data2[,4]=="weekend"),]
weekday_agg <- aggregate(steps~interval, data=weekday,FUN=mean)
weekend_agg <- aggregate(steps~interval, data=weekend,FUN=mean)
plot(weekday_agg,type="l", col="red")
lines(weekend_agg,type="l", col="blue")
legend(1, 200, c("weekday","weekend"), cex=0.8, col=c("red","blue"), lty=1:1);
```